Data Migration under a Changing Schema

### 2023-07-10 2022Migration/articleMigrationFACS.tex
First mathematical enforce rule

\usepackage{hyperref}
%\usepackage{multicol}
%\usepackage{footmisc}
%\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{mathtools}
%\usepackage{amsthm}
\usepackage[english]{babel}
%\usepackage[official,right]{eurosym}
\selectlanguage{english}
\hyphenation{ExecEngine}
%\newtheorem{lemma}{Lemma}

@@

   The notation $\schema_{\infsys}$ refers to the schema of information system $\infsys$.
   When clarity is needed, we write $\concepts_{\schema}$, $\rels_{\schema}$, and $\rules_{\schema}$
   for $\concepts$, $\rels$, and $\rules$ corresponding to $\schema$.

@@

   \begin{definition}[Event]
   Let $R \subseteq \Rels$ and $I \subseteq \Rels$ be sets of relations, and let $\infsys = \pair{\schema}{\dataset}$ and $\infsys' = \pair{\schema}{\dataset'}$ be information systems (so $\infsys$ and $\infsys'$ have the same schema), and:
      \begin{align}
      \triples_{\dataset} - (\Triple{\Atoms}{(R \cup I)}{\Atoms}) &= \triples_{\dataset'} - (\Triple{\Atoms}{(R \cup I)}{\Atoms})
   \label{eqn:eventUnchanged}\\
      \triples_{\dataset} - (\Triple{\Atoms}{R}{\Atoms}) &\subseteq \triples_{\dataset'} - (\Triple{\Atoms}{R}{\Atoms})
   \label{eqn:eventInsert}
   \end{align}
   Then we say that $\pair{\infsys}{\infsys'}$ is an event with scope $R$ inserting in $I$.
   We use the notation $\infsys \xrightarrow[I]{R} \infsys'$ to indicate this.
   \end{definition}

   Equation~\ref{eqn:eventUnchanged} states that the triples of those for relations in $I$ or $R$ are the only ones to have changed, and Equation~\ref{eqn:eventInsert} states that of the ones that are only in $I$, no triples were removed.
   If $I$ or $R$ is the empty set, we omit it from the arrow, so $\infsys \xrightarrow{R} \infsys'$ is a notation for $\infsys \xrightarrow[\emptyset]{R} \infsys'$.

@@

\subsection{Loading Data}
   Let us study how to migrate data from $\dataset$ to $\dataset'$.
   It would be tempting to state that a relation from the example system, say {\tt isPartOf}, could be copied by the following code fragment%
\footnote{The prefix {\tt old.} refers to concepts, relations, and rules from the existing system (section~\ref{sct:Example existing IS}).
The prefix {\tt new.} refers to items from the desired information system.}:
\begin{verbatim}
   ENFORCE new.isPartOf >: old.isPartOf
\end{verbatim}
   The relation `old.isPartOf' resides in the existing system and will not change after the MoT.

@@

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
   \begin{enumerate}
   \item We take a disjoint union of the relations: $\rels_\migrsys = \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}$
     where we define $\overleftarrow~$ and $\overrightarrow~$ to generate new names: $\overleftarrow{\declare{n}{x}{y}}=\declare{(old.n)}{x}{y}$ and $\overrightarrow{\declare{n}{x}{y}}=\declare{(new.n)}{x}{y}$,
     overloading the notations on sets pointwise: $\overleftarrow{X} = \{\overleftarrow{x}\mid x\in X\}$ and $\overrightarrow{X} = \{\overrightarrow{x}\mid x\in X\}$.
   \item We copy the relations 
   \end{enumerate}

>>>>

\usepackage{hyperref}
%\usepackage{multicol}
%\usepackage{footmisc}
%\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amssymb,stmaryrd}
\usepackage{mathrsfs}
\usepackage{mathtools}
%\usepackage{amsthm}
\usepackage[english]{babel}
%\usepackage[official,right]{eurosym}
\selectlanguage{english}
\hyphenation{ExecEngine}
%\newtheorem{lemma}{Lemma}

@@

   %The notation $\schema_{\infsys}$ refers to the schema of information system $\infsys$.
   When clarity is needed, we write $\concepts_{\schema}$, $\rels_{\schema}$, and $\rules_{\schema}$
   for $\concepts$, $\rels$, and $\rules$ corresponding to $\schema$.

@@

   \begin{definition}[Event]
   Let $R \subseteq \Rels$ and $I \subseteq \Rels$ be sets of relations, and let $\infsys' = \pair{\schema}{\dataset'}$ be an information system, and:
      \begin{align}
      \triples_{\dataset} - (\Triple{\Atoms}{(R \cup I)}{\Atoms}) &= \triples_{\dataset'} - (\Triple{\Atoms}{(R \cup I)}{\Atoms})
   \label{eqn:eventUnchanged}\\
      \triples_{\dataset} - (\Triple{\Atoms}{R}{\Atoms}) &\subseteq \triples_{\dataset'} - (\Triple{\Atoms}{R}{\Atoms})
   \label{eqn:eventInsert}
   \end{align}
   Then we say that $\pair{\pair{\schema}{\dataset}}{\infsys'}$ is an event with scope $R$ inserting in $I$.
   We use the notation $\pair{\schema}{\dataset} \xrightarrow[I]{R} \infsys'$ to indicate this.
   \end{definition}

   Equation~\ref{eqn:eventUnchanged} states that the triples of those for relations in $I$ or $R$ are the only ones to have changed, and Equation~\ref{eqn:eventInsert} states that of the ones that are only in $I$, no triples were removed.
   %If $I$ or $R$ is the empty set, we omit it from the arrow, so $\infsys \xrightarrow{R} \infsys'$ is a notation for $\infsys \xrightarrow[\emptyset]{R} \infsys'$.

@@

\subsection{Loading Data}\label{sec:loading}
   Let us study how to migrate data from $\dataset$ to $\dataset'$.
   It would be tempting to state that a relation from the example system, say {\tt isPartOf}, could be copied by the following code fragment%
\footnote{The prefix {\tt old.} refers to concepts, relations, and rules from the existing system (section~\ref{sct:Example existing IS}).
The prefix {\tt new.} refers to items from the desired information system.}:
\begin{verbatim}
   ENFORCE new.isPartOf >: old.isPartOf
\end{verbatim}
   The relation `old.isPartOf' resides in the existing system and will not change after the MoT.

@@

   Let $\pair{\dataset}{\schema}$ be the existing system, and let ${\schema'}$ be the schema of the desired system.
   Then we define the migration systems $\migrsys$ and the events as follows:
   \begin{enumerate}
   \item We take a disjoint union of the relations, defining $\overleftarrow~$ and $\overrightarrow~$ as (overloaded) notation to generate new names:
   \begin{align}
   \overleftarrow{\declare{n}{x}{y}}&=\declare{(\id{old}.n)}{x}{y}&\overleftarrow{X} &= \{\overleftarrow{x}\mid x\in X\}\\
   \overrightarrow{\declare{n}{x}{y}}&=\declare{(\id{new}.n)}{x}{y}& \overrightarrow{X} &= \{\overrightarrow{x}\mid x\in X\}\\
   \rels_1 &= \overleftarrow\rels_{\schema} \cup \overrightarrow\rels_{\schema'}
   \end{align}
   \item We create ENFORCE rules to copy equal relations, corresponding to (\ref{extraRelation}) and (\ref{difference}) in Section~\ref{sec:loading}:
   \begin{align}
   \widetilde{\declare{n}{x}{y}} &= \declare{(\id{copied}.n)}{x}{y}\\
   \rels_2 &= \{\widetilde{r}\mid r \in \rels_\schema \cap \rels_{\schema'}\}\\
   \triples_\migrsys &= \{\triple{a}{\overleftarrow{r}}{b} \mid \triple{a}{r}{b}\in\triples_\infsys\}\\
   \mathcal{E}_2 &= \{\overrightarrow{r}\mapsfrom (\overleftarrow{r} - \widetilde{r}) \mid \widetilde{r} \in \rels_2\} \cup \{\widetilde{r}\mapsfrom \overrightarrow{r} \mid \widetilde{r} \in \rels_2\}
   \end{align}
   \item 
   \end{enumerate}
